Abstractions for Multi-Sorted Substitutions

### 2024-03-14 paper/paper.tex
Update.

  The rest of this paper is structured as follows:
  Section~\ref{sec:syntax} introduces the multi-sorted syntax and compares it to the
  more common unsorted syntax.
  Section~\ref{sec:maps} introduces multi-sorted substitutions and renamings, and
  an abstraction to unify them.
  Section~\ref{sec:composition} introduces composition of multi-sorted
  substitions and renamings, and an abstraction to unify all four
  compositions.
  Section~\ref{sec:types} shows how to define types
  and type contexts for multi-sorted syntax.
  Section~\ref{sec:typing} shows how to define a multi-sorted typing relation,
  and an abstraction to unify type preservation lemmas for renamings and substitutions.
  % Section~\ref{sec:subjectreduction} introduces a small-step
  % operational semantics for System F and proves subject reduction.
  Section~\ref{sec:reflection} describes the class of object languages covered by
  our reflection algorithm.
  Section~\ref{sec:casestudies} discusses our case studies.
  Section~\ref{sec:related} discusses related work.
  Section~\ref{sec:conclusion} concludes.

@@

  \section{Types \& Contexts}
  \label{sec:types}
  % To define type contexts and a typing relation, we need to know
  % which sorts of terms play the role of types for other sorts of terms.
  % When modeling a typed language with multi-sorted syntax, both terms
  % and types are represented as terms, but with different sorts.
  % To define type contexts and a typing relation, we need to record for each term
  % from 
  % describe terms, and ocan be mapped to 
  % In a typed language,

@@

  \subsection{Type Contexts}
  Equipped with a notion of types, we define a type context as a
  function that maps variables to types of the same sort.

  As we want our framework to support dependent types, we allow a type returned
  by the type context to use all variables bound previously in the type context.
  This is achieved by the \ACode{drop-∈} function, which cuts off all
  sorts from \ACode{S} that are at smaller or equal positions than a variable \ACode{x}.
  Contexts \ACode{Γ} can be extended by a type \ACode{t} via \ACode{t \ADef{∷ᵗ} Γ}.

  % \LibCode\KContextHelper
  \noindent
  \begin{minipage}{1.0\linewidth}
    \begin{minipage}{0.5\linewidth}
      \LibCode\KContexts
      \vspace{3mm}
    \end{minipage}
    \begin{minipage}{0.5\linewidth}
      \LibCode\KContextExt
    \end{minipage}
  \end{minipage}

  To use a type returned from a type context, it needs to be
  weakened appropriately through the \ACode{\ADef{wk-telescope}} function:

  \noindent
  \begin{minipage}[t]{1.0\linewidth}
    \begin{minipage}[t]{0.45\linewidth}
      \LibCode\KContextLookupII
    \end{minipage}
    \begin{minipage}[t]{0.55\linewidth}
      \LibCode\KContextLookupI
    \end{minipage}
  \end{minipage}
  % \LibCode\KContextLookup

  We then define a variable typing as followed:
  \LibCode\KVariableTyping

  \section{Typing}
  \label{sec:typing}
  Now that we have a notion of types and type context, we are ready to define
  the multi-sorted typing relation for System F, which describes both
  typing and kinding:
  \AppCode\FTyping

>>>>

  The rest of this paper is structured as follows:
  Section~\ref{sec:syntax} introduces the multi-sorted syntax and compares it to the
  more common unsorted syntax.
  Section~\ref{sec:maps} introduces multi-sorted substitutions and renamings, and
  an abstraction to unify them.
  Section~\ref{sec:composition} introduces composition of multi-sorted
  substitions and renamings, and an abstraction to unify all four
  compositions.
  Section~\ref{sec:types} shows how to define types,
  type contexts, and typing relations,
  and presents an abstraction for unifying type preservation lemmas for renamings and substitutions.
  % Section~\ref{sec:subjectreduction} introduces a small-step
  % operational semantics for System F and proves subject reduction.
  Section~\ref{sec:reflection} describes the class of object languages covered by
  our reflection algorithm.
  Section~\ref{sec:casestudies} discusses our case studies.
  Section~\ref{sec:related} discusses related work.
  Section~\ref{sec:conclusion} concludes.

@@

  \section{Types \& Typing}
  \label{sec:types}
  \subsection{Types}
  % To define type contexts and a typing relation, we need to know
  % which sorts of terms play the role of types for other sorts of terms.
  % When modeling a typed language with multi-sorted syntax, both terms
  % and types are represented as terms, but with different sorts.
  % To define type contexts and a typing relation, we need to record for each term
  % from 
  % describe terms, and ocan be mapped to 
  % In a typed language,

@@

  \subsection{Type Contexts}
  \begin{figure}
    \centering
    % \LibCode\KContextHelper
    \begin{minipage}{1.0\linewidth}
      \begin{minipage}{0.57\linewidth}
        \LibCode\KContexts
        \LibCode\KContextLookupI
        \LibCode\KVariableTyping
      \end{minipage}
      \begin{minipage}{0.43\linewidth}
        \LibCode\KContextExt
        \LibCode\KContextLookupII
        \vspace{10.1mm}
      \end{minipage}
    \end{minipage}
    \caption{Type Contexts}
    \label{fig:type-contexts}
  \end{figure}
  Equipped with a notion of types, we are ready to define type contexts as shown
  in Figure~\ref{fig:type-contexts}.
  The type \ACode{\ADef{Ctx} S} describes type contexts with bindings for variables from \ACode{S}
  as functions that map variables to types of the same sort.
  As we want our framework to support dependent types, we allow a type
  in the context to use all variables bound previously in the context.
  This is achieved with \ACode{\ADef{drop-∈} x S}, which cuts off all
  sorts from \ACode{S} that are at smaller or equal positions than a variable \ACode{x}.
  The \ACode{\ADef{\_∷ᵗ\_}} function extends a context with a new binding.
  The \ACode{\ADef{lookup}} function looks up the type of a variable in a context
  and weakens it appropriately.
  Finally, a variable typing \ACode{Γ \ADef{∋} x \ADef{∶} t} is defined in terms of \ACode{\ADef{lookup}}.

  \subsection{Typing}
  \label{sec:types:typing}
  Now that we have a notion of types and type contexts, we are ready to define
  the multi-sorted typing relation for System F, which describes both
  typing and kinding:
  \AppCode\FTyping
